*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-21 Code/Manuscript_O2.Rmd
GAM results text edits.

## *Prochlorococcus marinus* growth responses to photoperiod, PAR and O~2~

@@

#### Generalized Additive Model (GAM) of MED4 growth {.unnumbered}

@@

Due to the complex interactions of light level and photoperiod, a GAM
was employed (Figure \@ref(fig:MED4GAM)) to model the growth trends
of MED4 in response to changing blue and red light PAR and photoperiod, across
the 3 different imposed oxygen concentrations. Under 250 µM O~2~, Figure
\@ref(fig:MED4GAM)A showed that the fastest growth rates were
achieved above 90 µmol photons m^-2^ s^-1^ and photoperiods greater than
12 h, indicated by the contour line labeled 0.62 d^-1^ (representing the
90^th^ percentile of achieved growth rate). Figure
\@ref(fig:MED4GAM)A also illustrated that MED4 exploited all light
levels as long as the photoperiods were longer than 4 h, albeit at
slower growth rate at lower PAR and shorter photoperiods. Under 25 µM
O~2~, MED4 grew, but required longer than 4 h photoperiod (Figure
\@ref(fig:MED4GAM)B). Interestingly, the GAM illustrated no
difference in growth rate when MED4 was grown under 25 µM or 250 µM O~2~
with photoperiods greater than 12 h. Figure \@ref(fig:MED4GAM)C
showed that MED4 did not grow under the tested conditions at 2.5 µM
O~2~.

@@

MED4GAM_cap <- glue("**A contour plot of a Generalized Additive Model (GAM) representing the chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MED4 grown under 660 nm (red) and 450 nm (blue) light.**",
                    "X-axis is photoperiod (h).",
                    "Y-axis is actinic Photosynethetic Active Radiation (PAR, µmol photons m^-2^ s^-1^).",
                    "**A.** represents the model under 250 µM of O~2~ and red light.",
                    "**B.** represents the model under 25 µM of O~2~ and red light.",
                    "**C.** represents the model under 250 µM of O~2~ and blue light.",
                    "**D.** represents the model under 25 µM of O~2~ and blue light.",
                    "Legends represent a colour gradient of growth rate from no growth (white) to 1.00 d^-1^ (dark red or dark blue).",
                    "Labeled contour lines indicate the 90%, 50%, and 10% quantiles for achieved growth rate.",

@@

#### Generalized Additive Model (GAM) of SS120 growth {.unnumbered}

@@

The GAM model in Figure \@ref(fig:SS120GAM) displays the growth
trends of SS120 in response to changing blue light PAR and photoperiod,
across the 3 different imposed oxygen concentrations. Under 250 µM O~2~,
Figure \@ref(fig:SS120GAM)A illustrated that the fastest growth
rates were achieved below PAR of 50 µmol photons m^-2^ s^-1^ and
photoperiods between 8 and 12 h, indicated by the contour line labeled
0.29 d^-1^ (representing the 90^th^ percentile of achieved growth rate).
Figure \@ref(fig:SS120GAM)A also showed that SS120 does not grow at
PAR above 90 µmol photons m^-2^ s^-1^ and the growth rate are slower as
photoperiods go above 12 or below 8 h. Under 25 µM O~2~, SS120 grew
under a photoperiod of 8 h and exploited all light levels and grew under
a 12 h photoperiod but at PAR of 50 µmol photons m^-2^ s^-1^. The
fastest growth rate achieved was at a higher PAR of \~100 µmol photons
m^-2^ s^-1^, indicated by the contour line labeled 0.5 d^-1^
(representing the 90^th^ percentile of achieved growth rate), out pacing
the 90^th^ percentile fastest growth rates under 250 µM O~2~. Figure
\@ref(fig:SS120GAM)C showed that SS120 did not grow under the tested
conditions at 2.5 µM O~2~.

@@

SS120GAM_cap <- glue("**Contour plot of a Generalized Additive Model (GAM) representing the chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 grown under 660 nm (red) and 450 nm (blue) light.**",
                    "X-axis is photoperiod (h).",
                    "Y-axis is actinic Photosynethetic Active Radiation (PAR, µmol photons m^-2^ s^-1^).",
                    "**A.** represents the model under 250 µM of O~2~ and red light.",
                    "**B.** represents the model under 25 µM of O~2~ and red light.",
                    "**C.** represents the model under 250 µM of O~2~ and blue light.",
                    "**D.** represents the model under 25 µM of O~2~ and blue light.",
                    "Legends represent a colour gradient of growth rate from no growth (white) to 1.00 d^-1^ (dark red or dark blue).",
                    "Labeled contour lines indicate the 90%, 50%, and 10% quantiles for achieved growth rate.",

@@

#### Generalized Additive Model (GAM) of MIT9313 growth {.unnumbered}

The GAM model in Figure \@ref(fig:MIT9313GAM) displays the growth
trends of MIT9313 in response to changing blue light PAR and
photoperiod, across the 3 different imposed oxygen concentrations. Under
250 µM O~2~, Figure \@ref(fig:MIT9313GAM)A illustrated that the
fastest growth rates were achieved between PAR of 30 µmol photons m^-2^
s^-1^ and 50 µmol photons m^-2^ s^-1^ and photoperiods longer than 8 h,
indicated by the contour line labeled 0.51 d^-1^ (representing the
90^th^ percentile of achieved growth rates). Figure
\@ref(fig:MIT9313GAM)A also illustrated that growth rate increases
with longer photoperiods, as long as the PAR levels are below 50 µmol
photons m^-2^ s^-1^ and that MIT9313 does not grow when exposed to PAR
levels above 90 µmol photons m^-2^ s^-1^. In contrast to the 250 µM O~2~
growth experiment, MIT9313 was able to exploit all actinic PAR levels
(Figure \@ref(fig:MIT9313GAM)B). Interestingly, the growth rate was
similar across all PAR levels between 4 and 12 h photoperiods, indicated
by the contour line labeled 0.29 d^-1^ (representing the 10^th^
percentile of achieved growth rate) and 0.44 d^-1^ (representing the
50^th^ percentile of achieved growth rate) (Figure
\@ref(fig:MIT9313GAM)B). However, for photoperiods exceeding 12 h,
the growth rate showed a decline at higher PAR levels but showed the
overall fastest growth rate achieved was at PAR levels less than 100
µmol photons m^-2^ s^-1^, indicated by the contour line labeled 0.69
d^-1^ (representing the 90^th^ percentile of achieved growth rate)
(Figure \@ref(fig:MIT9313GAM)B). Figure \@ref(fig:MIT9313GAM)C
showed that MIT9313 is capable of growing under 2.5 µM O~2~ as long as
the photoperiod is between 4 and 8 h and between 50 to 150 µmol photons
m^-2^ s^-1^ PAR levels (Figure \@ref(fig:MIT9313GAM)C).

@@

MIT9313GAM_cap <- glue("**Contour plot of a Generalized Additive Model (GAM) representing the chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 grown under 660 nm (red) and 450 nm (blue) light.**",
                    "X-axis is photoperiod (h).",
                    "Y-axis is actinic Photosynthetically Active Radiation (PAR, µmol photons m^-2^ s^-1^).",
                    "**A.** represents the model under 250 µM of O~2~ and red light.",
                    "**B.** represents the model under 25 µM of O~2~ and red light.",
                    "**C.** represents the model under 2.5 µM of O~2~ and red light.",
                    "**D.** represents the model under 250 µM of O~2~ and blue light.",
                    "**E.** represents the model under 25 µM of O~2~ and blue light.",
                    "**F.** represents the model under 2.5 µM of O~2~ and blue light.",
                    "Legends represent a colour gradient of growth rate from no growth (white) to 1.00 d^-1^ (dark red or dark blue).",
                    "Labeled contour lines indicate the 90%, 50%, and 10% quantiles for achieved growth rate.",

>>>>

## *Prochlorococcus marinus* growth responses to photoperiod, PAR, spectral bands,  and O~2~

@@

## Generalized Additive Models (GAM) of *Prochlorococcus marinus* growth responses to photoperiod, PAR, spectal bands and O~2~   {.unnumbered}

@@

We employed GAM to summarize and visualize th complex responses of growth to the interactions of peak PAR, oxygen, photoperiod and blue vs. red spectral bands

### MED4 Clade HLI
The GAM model in Figure \@ref(fig:MED4GAM) shows MED4 growth in response to red (A) or blue (B) peak PAR and photoperiod across 2 imposed oxygen concentrations.  Under 250 µM O~2~ MED4 achieved fastest growth rates above peak blue light of ~180 µmol photons m^-2^ s^-1^, and the longest photoperiod of 16, indicated by the 0.69 d^-1^ contour line representing the
90^th^ percentile of maximum achieved growth rate.  Growth decreased with decreasing photoperiod and peak PAR.  Under red light growth was generally slower but the pattern of growth responses to photoperiod and PAR waas similar.  We again note the exclusion of MED4 from growth under 4 h photoperiod. Under 25 µM O~2~ MED4 showed similar growth responses, but was excluded from both 4 and 8 h photoperiods.

@@

MED4GAM_cap <- glue("**A contour plot of a Generalized Additive Model (GAM) representing the chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MED4 grown under 660 nm (red) or 450 nm (blue) light.**",
                    "X-axis is photoperiod (h).",
                    "Y-axis is actinic Photosynethetic Active Radiation (PAR, µmol photons m^-2^ s^-1^).",
                    "**A.** represents the model under 250 µM of O~2~ and red light.",
                    "**B.** represents the model under 25 µM of O~2~ and red light.",
                    "**C.** represents the model under 250 µM of O~2~ and blue light.",
                    "**D.** represents the model under 25 µM of O~2~ and blue light.",
                    "Legends represent a colour gradient of growth rate from no growth (white) to 1.00 d^-1^ (dark red or dark blue).",
                    "Labeled contour lines indicate the 90%, 50%, and 10% quantiles for achieved growth rate.",

@@

### SS120 Clade Clade LLIII  {.unnumbered}

@@

The GAM model in Figure \@ref(fig:SS120GAM) shows growth
responses of SS120 red (A) or blue (B) peak PAR and photoperiod,
across the 2 imposed oxygen concentrations. Under 250 µM O~2~,
Figure \@ref(fig:SS120GAM)A showed highest growth
rates achieved below PAR of 50 µmol photons m^-2^ s^-1^ and
photoperiods between 8 and 12 h, indicated by the contour line labeled
0.29 d^-1^ (representing the 90^th^ percentile of achieved growth rate). Note the disjunct regions of the plots, resulting from variable growth successes of SS120 under 250 µM O~2~.  Growth rate patterns under red light were similar, although somehwat faster.
In contrast, under 25 µM O~2~ and a photoperiod of 8 h SS120 exploited all blue peak PAR levels, achieving faster growth ratesat a higher PAR of \~100 µmol photons
m^-2^ s^-1^, indicated by the contour line labeled 0.5 d^-1^
(representing the 90^th^ percentile of achieved growth rate), out pacing
the 90^th^ percentile fastest growth rates under 250 µM O~2~. Under red light and 25 µM O~2~ light SS120 grew across most conditions of peak PAR and photoperiod, achieving fastest growhth under long photoperiods and peak PAR below 50 \~100 µmol photons
m^-2^ s^-1^.  Thus, the designation of SS120 as a LL strain is dependent upon the [O~2]. SS120 did not, however,  grow under tested conditions at 2.5 µM O~2~.

@@

SS120GAM_cap <- glue("**Contour plot of a Generalized Additive Model (GAM) representing the chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* SS120 grown under 660 nm (red) or 450 nm (blue) light.**",
                    "X-axis is photoperiod (h).",
                    "Y-axis is actinic Photosynethetic Active Radiation (PAR, µmol photons m^-2^ s^-1^).",
                    "**A.** represents the model under 250 µM of O~2~ and red light.",
                    "**B.** represents the model under 25 µM of O~2~ and red light.",
                    "**C.** represents the model under 250 µM of O~2~ and blue light.",
                    "**D.** represents the model under 25 µM of O~2~ and blue light.",
                    "Legends represent a colour gradient of growth rate from no growth (white) to 1.00 d^-1^ (dark red or dark blue).",
                    "Labeled contour lines indicate the 90%, 50%, and 10% quantiles for achieved growth rate.",

@@

The GAM model in Figure \@ref(fig:MIT9313GAM) shows MIT9313 growth responses to red (A) or blue (B) peak PAR and photoperiod. Under 250 µM O~2~,  Figure \@ref(fig:MIT9313GAM)A shows MIT9313 achieves
fastest growth rates between blue peak PAR of 30 µmol photons m^-2^
s^-1^ and 50 µmol photons m^-2^ s^-1^ and photoperiods longer than 8 h,
indicated by the contour line labeled 0.51 d^-1^ representing the
90^th^ percentile of achieved growth rates. Figure
\@ref(fig:MIT9313GAM)A also shows that growth rate increases
with longer photoperiods, as long as the blue peak PAR levels remain below 50 µmol
photons m^-2^ s^-1^.  In contrast, under red light and 250 µM O~2~ MIT9313 grows faster while exploiting higher peak PAR and longer photoperiods.  Thus the designation of MIT9313 as a LL clade is dependent upon [O2].
(Figure \@ref(fig:MIT9313GAM)B).  Figure \@ref(fig:MIT9313GAM)C
shows that MIT9313 maintains growth even under 2.5 µM O~2~, under photoperiods between 4 and 8 h and peak blue PAR between 50 to 10 µmol photons
m^-2^ s^-1^ PAR (Figure \@ref(fig:MIT9313GAM)C).

@@

MIT9313GAM_cap <- glue("**Contour plot of a Generalized Additive Model (GAM) representing the chlorophyll specific growth rate (d^-1^) for *Prochlorococcus marinus* MIT9313 grown under 660 nm (red) or 450 nm (blue) light.**",
                    "X-axis is photoperiod (h).",
                    "Y-axis is actinic Photosynthetically Active Radiation (PAR, µmol photons m^-2^ s^-1^).",
                    "**A.** represents the model under 250 µM of O~2~ and red light.",
                    "**B.** represents the model under 25 µM of O~2~ and red light.",
                    "**C.** represents the model under 2.5 µM of O~2~ and red light.",
                    "**D.** represents the model under 250 µM of O~2~ and blue light.",
                    "**E.** represents the model under 25 µM of O~2~ and blue light.",
                    "**F.** represents the model under 2.5 µM of O~2~ and blue light.",
                    "Legends represent a colour gradient of growth rate from no growth (white) to 1.00 d^-1^ (dark red or dark blue).",
                    "Labeled contour lines indicate the 90%, 50%, and 10% quantiles for achieved growth rate.",
